A deeper look into bug fixes: patterns, replacements, deletions, and additions

### 2016-02-17 MSRChallenge2016.tex
Revise results section and threats to validity
* Slightly changed the story in results section
* Added threats to validity section
* Integrate CPW's replacement results
* Added some todos (mostly for CPW)
* Fixed a few typos
* Missing conclusion

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Mauricio Soto\\
       \affaddr{School of Computer Science}\\
       \affaddr{Carnegie Mellon University}\\
       \affaddr{Pittsburgh, PA}\\
       \email{mauriciosoto@cmu.edu}
       % 3rd. author
\alignauthor Ferdian Thung\\
       \affaddr{University}\\
       \affaddr{City}\\
       \email{email@domain.com}
% 2nd. author
\alignauthor Chu Pan Wong   \\
       \affaddr{Carnegie Mellon University}\\
       \affaddr{Pittsburgh, PA}\\
       \email{email@domain.com}       
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor
Claire Le Goues\\
       \affaddr{School of Computer Science}\\
       \affaddr{Carnegie Mellon University}\\
       \affaddr{Pittsburgh, PA}\\
       \email{clegoues@cs.cmu.edu}
% 5th. author
\alignauthor
David Lo\\
       \affaddr{University}\\
       \affaddr{City}\\
       \email{author@institution.edu}       
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{February 2, 2016}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

\section{Related Work}
We take as reference the study performed by Dongsun Kim et al.~\cite{kim2013} in which they look for the most common ways in which programmers patch bugs in software. The
researchers developed a variation of the tool Genprog~\cite{weimer2009,legoues2012}
with several different templates resembling patters programmers use to patch
bugs, and then perform an empirical study to evaluate which patches do human programmers prefer.

@@

The above patterns only cover 8 out of 10 bug fixing patterns in PAR. The other two patterns are: (1) checking that an object is an instance of a target class before casting it; and (2) replacing expression in an if condition by another expression in the same scope. Pattern (1) cannot be analyzed because Boa does not support {\em instanceof} expression. Similarly, pattern (2) requires us to track scope between pre-fix and post-fix version, which is not supported by Boa. Implementing them ourselves might be possible, but it might require too much effort than necessary. Investigating 8 out of 10 patterns should give us a good idea on how common the patterns actually are.

@@

\begin{enumerate}
\item {\bf AMP Pattern}\\
  To detect AMP pattern, for both pre-fix and post-fix version of a buggy file,
  we create a custom method call signature. This signature contains method name,
  literal parameter, and variable parameter. we consider parameter containing
  custom expression as OTHER. This is due to inexistence of function that can
  print Abstract Syntax Tree (AST) back to source code. we discard method
  signatures that appear both in pre-fix and post-fix version. We then find
  whether there exists method with the same name and the same number of
  parameters, but different parameter signatures in pre-fix and post-fix
  version. If it exists, we consider that AMP pattern is found.

@@

\todo{What is the dataset?}

This section investigates how do developer fix bugs from three different
perspectives. To begin with, we show some characteristics of real-world bug
fixes.  Afterwards, we look deeper to find out what kinds of statements are
usually added, modified or deleted by human developers to fix bugs. Lastly, we
show how common PAR's fixing templates are in our dataset.

\subsection{Characteristics}

This subsection shows a bird's-eye view of real-world bug fixes at
ultra-large-scale. We did our analysis on the \emph{2015 September/Github}
dataset on Boa. This dataset includes $7,830,023$ Java projects with
$23,229,406$ revisions. Among all the revisions, $4,590,679$ revisions are
identified by Boa as bug fixing revisions. Our analysis in this subsection
focuses on these bug fixing revisions. We hope that this population
is big enough to reflect the bug fixing practice.

To begin with, we are interested in:

@@

To answer this question, we identified all the bug fixing revisions in our
dataset and count the number of changed files. In total, $52,052,571$ files got
changed. Thus, on average each bug fixing revision changes 11.3 files.  This
number is surprisingly high because most automatic program repair techniques
assume that bugs are local in most cases. There could be several reasons behind
this number. It could be the case that other software artifacts (e.g.
documentation) are updated after bug fixing. It could also be the case that bug
fixing commit also contains changes that are not related to fixing bug (e.g. new
features, refactoring). To better understand why so many files are changed, we
continue to ask:

@@

In Boa, each changed file is represented by a \emph{ChangedFile} node, and it
has a field called \emph{FileKind}, describing the type of each file.
Table~\ref{tbl:fileType} shows the file types that Boa could identify. To answer
this question, we iterate all the changed files in each bug fixing revision and
count the total number for each file type. Results are shown in
Table~\ref{tbl:fileType}.

\begin{table}
\centering
  \begin{tabular}{| c | r | r |}
  \hline
  File Kind & Total Number & Average \\ \hline \hline
  BINARY & 752,945 & 0.16 \\ \hline
  SOURCE\_JAVA\_ERROR & 2,073,558 & 0.45 \\ \hline
  SOURCE\_JAVA\_JLS2 & 2,607,413 & 0.57 \\ \hline
  SOURCE\_JAVA\_JLS3 & 15,748,967 & 3.43 \\ \hline
  SOURCE\_JAVA\_JLS4 & 83,798 & 0.02 \\ \hline
  TEXT & 541,023 & 0.12 \\ \hline
  XML & 6,818,299 & 1.49 \\ \hline
  UNKNOWN & 23426568 & 5.10 \\ \hline
  \end{tabular}
  \caption{File Types of Changed Files}
  \label{tbl:fileType}
\end{table}

@@

This research question is interesting because most automatic program repair
approaches could not create new classes/methods/fields/variables to fix bugs.
Most approaches guess and synthesize the patches based on existing source code.

@@

\todo{Is there room/call to include any element of Chu Pan's replacement
  analysis in here?}

\subsection{Frequency of PAR Bug Fixing Pattern}\label{sec:freqfixpattern}
After running the bug pattern detection approach on Boa's dataset, I obtained the result shown in Table~\ref{tab:freqpattern}. The result shows that AMC, MSM, COM, CBC, IAO, ANC, AOB, and COB pattern is found on 1.95\%, 1.69\%, 0.58\%, 4.23\%, 2.82\%, 2.90\%, 0.28\%, and 0.57\% of buggy files. We can see that the most common pattern among these patterns is CBC and the least common pattern is AOB. If we assume that these patterns never appear together in the dataset, these patterns cover 14.78\% of buggy files. Kim et al. mentioned that the patterns that they use cover almost 30\% of real patches~\cite{kim2013}. Although we only detect 8 out of 10 bug fixing patterns, it is hard to believe that the remaining two patterns cover 15\% of buggy files, especially 14.78\% is a number we get when we assume that the patterns that we detect never appear together. Moreover, the percentage of each pattern is based on imprecise approximation which is likely to represent an upper bound rather than a lower bound. This suggests that common bug fixing patterns might not be as common as it appeared to be.

\subsection{Data Analysis}
Our queries were ran on the dump of Github of September 2015. In these real life
projects we found 46,301,429 files being fixed. In these files, there is a total
of 517,281 statement kinds being replaced, and 511,198 replacing these
statements kinds.

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[scale=0.5]{g1.png}
  \caption{Most common replacements}
\end{figure}

@@

\begin{figure}
  \centering
\includegraphics[scale=0.5]{g2.png}
\caption{Least common replacements}
\end{figure}

@@

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns

>>>>

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Mauricio Soto\\
       \affaddr{School of Computer Science}\\
       \affaddr{Carnegie Mellon University}\\
       \affaddr{Pittsburgh, PA}\\
       \email{mauriciosoto@cmu.edu}
       % 3rd. author
\alignauthor Ferdian Thung\\
       \affaddr{University}\\
       \affaddr{City}\\
       \email{email@domain.com}
% 2nd. author
\alignauthor Chu-Pan Wong   \\
       \affaddr{Carnegie Mellon University}\\
       \affaddr{Pittsburgh, PA}\\
       \email{email@domain.com}       
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor
Claire Le Goues\\
       \affaddr{School of Computer Science}\\
       \affaddr{Carnegie Mellon University}\\
       \affaddr{Pittsburgh, PA}\\
       \email{clegoues@cs.cmu.edu}
% 5th. author
\alignauthor
David Lo\\
       \affaddr{University}\\
       \affaddr{City}\\
       \email{author@institution.edu}       
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{February 2, 2016}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

\section{Related Work}
We take as reference the study performed by Dongsun Kim et al.~\cite{kim2013} in which they look for the most common ways in which programmers patch bugs in software. The
researchers developed a variation of the tool Genprog~\cite{weimer2009,legoues2012}
with several different templates resembling patterns programmers use to patch
bugs, and then performed an empirical study to evaluate which patches do human programmers prefer.

@@

The above patterns only cover 8 \todo{inconsistent with intro} out of 10 bug fixing patterns in PAR. The other two patterns are: (1) checking that an object is an instance of a target class before casting it; and (2) replacing expression in an if condition by another expression in the same scope. Pattern (1) cannot be analyzed because Boa does not support {\em instanceof} expression. Similarly, pattern (2) requires us to track scope between pre-fix and post-fix version, which is not supported by Boa. Implementing them ourselves might be possible, but it might require too much effort than necessary. Investigating 8 out of 10 patterns should give us a good idea on how common the patterns actually are.

@@

\begin{enumerate}
\item {\bf AMP Pattern}\\
  To detect AMP pattern, for both pre-fix and post-fix version of a buggy file,
  we create a custom method call signature. This signature contains method name,
  literal parameter, and variable parameter. we consider parameter containing
  custom expression as OTHER. This is due to inexistence of function that can
  print Abstract Syntax Tree (AST) back to source code in Boa. We discard method
  signatures that appear both in pre-fix and post-fix version. We then find
  whether there exists method with the same name and the same number of
  parameters, but different parameter signatures in pre-fix and post-fix
  version. If it exists, we consider that AMP pattern is found.

@@

% (This is the old version)
% This section investigates how do developer fix bugs from three different
% perspectives. To begin with, we show some characteristics of real-world bug
% fixes.  Afterwards, we look deeper to find out what kinds of statements are
% usually added, modified or deleted by human developers to fix bugs. Lastly, we
% show how common PAR's fixing templates are in our dataset.

This section shows our analysis results. To begin with, we discuss the dataset
we are using, and some characteristics of real-world bug fixes. Afterwards, we
provide insights for two main categories of automatic program repair approaches.
On the one hand, many approaches \todo{citation} rely on mutation operators to
modify the source code, but we have limited knowledge about what kinds of
statements are usually modified. Our analysis results will fill in this
knowledge gap. On the other hand, some automatic program repair approaches
\todo{citation} search for specific patterns and modify source code according to
predefined templates. Our results provide evidence of how common those
patterns are in practice.

\subsection{Dataset and Characteristics}

We did our analysis on the \emph{2015 September/Github} dataset on Boa. This
dataset includes $7,830,023$ Java projects with $23,229,406$ revisions. Among
all the revisions, $4,590,679$ revisions are identified by Boa as bug fixing
revisions. We hope that this population is big enough to reflect the bug fixing
practice.

The rest of this subsection shows some characteristics of real-world bug fixes.
To begin with, we are interested in:

@@

To answer this question, we identify all the bug fixing revisions in our dataset
and count the number of changed files. A file was changed if (1) it was a new
file; (2) or it was modified; (3) or it was deleted. In total, $52,052,571$
files got changed. Thus, on average each bug fixing revision changes 11.3 files.
This number is surprisingly high because most automatic program repair
techniques assume that bugs are local in most cases. There could be several
reasons behind this number. It could be the case that other software artifacts
(e.g.  documentation) are updated after bug fixing. It could also be the case
that bug fixing commit also contains changes that are not related to fixing bug
(e.g. refactoring, adding new features by creating new files). To better
understand why so many files are changed, we continue to ask:

@@

In Boa, each changed file is represented by a \emph{ChangedFile} node, and it
has a field called \emph{FileKind}, describing the type of each file. As
mentioned earlier, a \emph{ChangedFile} could be either a new file or an
existing file (if the bug fixing revision modified or deleted the file).
Table~\ref{tbl:fileType} shows the file types that Boa could identify. To answer
this question, we iterate all the changed files in each bug fixing revision and
count the total number for each file type. Results are shown in
Table~\ref{tbl:fileType}.

\begin{table}
\centering
  \begin{tabular}{| c | r | r |}
  \hline
  File Kind & Total Number & Average \\ \hline \hline
  BINARY & 752,945 & 0.16 \\ \hline
  SOURCE\_JAVA\_ERROR & 2,073,558 & 0.45 \\ \hline
  SOURCE\_JAVA\_JLS2 & 2,607,413 & 0.57 \\ \hline
  SOURCE\_JAVA\_JLS3 & 15,748,967 & 3.43 \\ \hline
  SOURCE\_JAVA\_JLS4 & 83,798 & 0.02 \\ \hline
  TEXT & 541,023 & 0.12 \\ \hline
  XML & 6,818,299 & 1.49 \\ \hline
  UNKNOWN & 23,426,568 & 5.10 \\ \hline
  \end{tabular}
  \caption{File Types of Changed Files}
  \label{tbl:fileType}
\end{table}

@@

This research question is interesting because most automatic program repair
approaches often do not create new classes/methods/fields/variables to fix bugs.
Most approaches guess and synthesize the patches based on existing source code.

@@

% \todo{Is there room/call to include any element of Chu Pan's replacement
%   analysis in here?}

\subsection{What changes are common in real-world bug fixes?}

Most automatic program repair approaches mutate source code using predefined
mutation operators, which represent different ways of modifying source code. In
order to generate patches that make more sense to human developers, it would be
beneficial to first investigate how developers mutate source code to fix bugs.

For each bug fixing revision $r$ and each statement kind $k$, we count how many
statements of kind $k$ are there in revision $r$ and $r-1$ ($r-1$ means the
revision before revision $r$), and then compare the two numbers. For example, if
the number of if statements in revision $r$ is more than the number in revision
$r-1$, we say bug fixing revision $r$ introduce new if statements. Note that
``introducing new if statements'' does not necessarily mean human developers
introduce them to fix bugs. However, our hope is that this analysis could still
provide us some hints, given the large number of bug fixing revisions we are
analyzing.

\begin{figure*}[!t] \centering
	\includegraphics[width=0.95\textwidth]{StmtType} \caption{For each kind
	of statement, how many bug fixing revisions add/delete statements of the
same kind?} \label{fig:StmtType} \end{figure*}

From Figure~\ref{fig:StmtType} \todo{fix the small fonts in this graph}, we can
see that expression, if, return, for and try catch statements are added most
often. On the other hand, the same types of statements are also deleted most
often. This finding indicates that most bugs were fixed by changing the control
flow of the program using if, return, for, while, and try catch statements.

For each bug fixing revision, the results discussed above consider both newly
added files and existing files. Since most automatic program repair approaches
do not generate new files to fix bugs, the rest of this section only considers
files that were already in the repositories when developers fixed bugs. Our
analysis shows that there are 46,301,429 of them. In these files, there is a
total of 517,281 statement kinds being replaced, and 511,198 replacing these
statements kinds.

@@

\begin{figure}[!t]
  \centering
      \includegraphics[scale=0.5]{g1.png}
  \caption{Most common replacements}
\end{figure}

@@

\begin{figure}[!t]
  \centering
\includegraphics[scale=0.5]{g2.png}
\caption{Least common replacements}
\end{figure}

\subsection{How common are the fixing templates of PAR?}\label{sec:freqfixpattern}
After running the bug pattern detection approach on Boa's dataset, I obtained the result shown in Table~\ref{tab:freqpattern}. The result shows that AMC, MSM, COM, CBC, IAO, ANC, AOB, and COB pattern is found on 1.95\%, 1.69\%, 0.58\%, 4.23\%, 2.82\%, 2.90\%, 0.28\%, and 0.57\% of buggy files. We can see that the most common pattern among these patterns is CBC and the least common pattern is AOB. If we assume that these patterns never appear together in the dataset, these patterns cover 14.78\% of buggy files. Kim et al. mentioned that the patterns that they use cover almost 30\% of real patches~\cite{kim2013}. Although we only detect 8 out of 10 bug fixing patterns, it is hard to believe that the remaining two patterns cover 15\% of buggy files, especially 14.78\% is a number we get when we assume that the patterns that we detect never appear together. Moreover, the percentage of each pattern is based on imprecise approximation which is likely to represent an upper bound rather than a lower bound. This suggests that common bug fixing patterns might not be as common as it appeared to be.

\section{Threats to Validity}

Since we are using a DSL provided by Boa to implement our analysis, the
correctness of our analysis depends on our programs as well as Boa itself. To
prevent errors from our side, we open source our implementation on
Github~\footnote{\url{https://github.com/}}.\todo{update
this}

Some threats to our study are due to the limitation of Boa. For example, we rely
on Boa to identify bug fixing revisions, but it is unclear whether those
identified revisions are actually related to bug fixing. Chances are that there
exist some other bug fixing revisions, but Boa could not identify them. It is
also possible that bug fixing revisions identified by Boa actually include other
code changes, such as implementing new features, refactoring the code base.
After all, identifying bug fixing revisions from revision history is still an
open problem. Another big limitation of Boa is that there is no easy to get a
precise statement-level diff between two revisions. Thus, our template matching
and analysis of code changes (by counting each statement kind) could be
inaccurate. However, our results should provide upper bounds for the questions
we investigated.

@@

\todo{Fix the BibTeX}

\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
